---
title: "Transit Accessibility Maps in R"
subtitle: "H3 Clustering, Routing APIs, and IDW Isolines"
author: "Aleksei Prishchepo"
date: "2026-05-18"
number-sections: true
format:
  titlepage-pdf:
    titlepage: formal 
    titlepage-logo: "image.png"
    titlepage-theme:
      elements: ["\\titleblock", "Practitioner walkthrough of building a transit accessibility pipeline in R — covering H3 hexagonal clustering, resilient routing calculation loops, and IDW isoline generation over an OSM base tile.", "\\vfill", "\\logoblock", "\\authorblock", "\\dateblock"]
      page-align: "center"
      title-style: "plain"
      title-fontstyle: ["Huge", "textbf"]
      title-space-after: "1.5cm"
      subtitle-fontstyle: "LARGE"
      title-subtitle-space-between: "2\\baselineskip"
      author-style: "plain"
      author-sep: "newline"
      author-fontstyle: "Large"
      author-space-after: "2\\baselineskip"
      affiliation-style: "numbered-list-with-correspondence"
      affiliation-fontstyle: "large"
      affiliation-space-after: "0pt"
      footer-style: "plain"
      footer-fontstyle: ["Large", "textsc"]
      footer-space-after: "0pt"
      logo-size: "0.8\\textwidth"
      logo-space-after: "1cm"
    documentclass: scrartcl
    toc: true
    toc-title: Contents
    toc-depth: 1
    number-sections: true
    number-depth: 1
    papersize: letter
    margin-bottom: 1.2in
    fig-pos: 'H'
    include-in-header: 
       text: |
         \usepackage{fvextra}
         \DefineVerbatimEnvironment{Highlighting}{Verbatim}{breaklines,commandchars=\\\{\}}
         \DefineVerbatimEnvironment{OutputCode}{Verbatim}{breaklines,commandchars=\\\{\}}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
library(tidyverse)
library(kableExtra)
```

## Introduction

The map on the cover of this article shows transit travel times from 189
hotel clusters to 50 cultural and nightlife destinations across Berlin.
Each hexagon is shaded by average travel time; the isolines mark
5-minute intervals; the red circles are 1, 3, and 5 km from the
geographic centre. The point it makes is geographic: transit
accessibility and physical proximity are different things.

This article is about how that map was built. The motivating problem was
a [hotel search
analysis](https://frequentist.org/posts/20260207-transit-time-hotel-search/)
— ranking hotels by transit time to a traveller's specific points of
interest rather than by distance from the city centre. But the pipeline
generalises to any problem where you need to measure accessibility
between a set of origins and destinations at city scale: venue analysis,
urban planning, catchment area modelling.

Three engineering decisions shape the structure. First, how to reduce
22,750 potential routing requests to something tractable without losing
spatial resolution — that's the H3 clustering section. Second, how to
run thousands of API calls in a loop that doesn't die on the network
error and doesn't create a memory bottleneck as results accumulate —
that's the routing loop section. Third, how to turn a sparse set of
point measurements into a readable continuous surface with labelled
isolines over a basemap — that's the visualisation section.

The code is in R. The libraries doing the work are `h3jsr` for hexagonal
indexing, `gmapsdistance` for routing, `gstat` and `stars` for IDW
interpolation, and `metR` and `ggspatial` for the final map composition.

## Load Data

The source data comes from the OpenStreetMap extract for Berlin via
[Freie Universität
Berlin](https://www.geo.fu-berlin.de/en/v/soga-r/Advances-statistics/Spatial-Point-Patterns/Berlin-City-Data/index.html),
containing features such as museums, theaters, parks, and other
attractions.

```{r load-data}
#| eval: false
temp <- tempfile()
download_url <- "https://userpage.fu-berlin.de/soga/data/raw-data/spatial/"
zipfile <- "berlin-latest-free.shp.zip"

## download the file
download.file(paste0(download_url, zipfile), temp, mode = "wb")
## unzip the file(s)
unzip(temp, exdir = "data")

## close file connection
unlink(temp)
```

```{r load-shapefile}
library(sf)

berlin_features <- st_read("data/gis_osm_pois_free_1.shp", quiet = TRUE)
berlin_features |> glimpse()
```

## Explore POI Data

The OSM extract contains dozens of feature classes. We need to narrow to
the ones that represent meaningful destinations for a visitor: cultural
venues, nightlife, landmarks, and exclude noise like parking lots and
post boxes.

```{r features-classes}
#| class-output: OutputCode

berlin_features$fclass |> unique() |> sort() |>
  paste(collapse = ", ") |>  cat()
```

```{r poi-filtering}
poi_classes <- c("archaeological", "attraction", "monument",
  "museum", "nightclub", "observation_tower", "ruins", 
  "graveyard", "stadium", "theatre", "zoo"
)

berlin_pois <- berlin_features |> filter(fclass %in% poi_classes)

berlin_pois |> glimpse()
```

```{r fig-poi-map, message=FALSE}
library(ggplot2)
library(sf)
library(ggspatial)

berlin_pois <- berlin_pois |> mutate(
  loc = st_coordinates(geometry),
  coord = map2(loc[,2], loc[,1], ~c(.x, .y))
) |> select(-loc) |> as.data.frame()

berlin_pois$lng <- berlin_pois$coord |> map_dbl(2)
berlin_pois$lat <- berlin_pois$coord |> map_dbl(1)

berlin_pois_sf <- berlin_pois |>
  sample_n(100) |>
  st_as_sf(coords = c("lng", "lat"), crs = 4326)

ggplot() +
  annotation_map_tile(type = "osm", zoom = 12) +
  geom_sf(data = st_transform(berlin_pois_sf, 3857),
          color = "purple",
          size = 2) +
  theme_minimal() +
  theme(
    axis.text = element_blank(),
    axis.title = element_blank()
  ) +
  labs(
    title = "Berlin POIs by Category",
    subtitle = "Cultural and nightlife locations"
  )
```

With all 11 POI classes included, the points are scattered across the
map forming no visible clusters. The next step is to rank by engagement
and trim to the destinations that are popular among travellers.

## Add Reviews Data to POIs

Google Places ratings serve as a proxy for a POI's relevance: places
with more reviews have been visited by more people. It worth noting that
ratings follow a heavily right-skewed distribution. The Berlin Zoo has
hundreds of thousands of reviews; a small neighbourhood museum might
have 800. Filtering naively by the top 50 on raw count produces a list
dominated by mass-market tourist attractions, which are already
well-served by central hotels. For the scenario-based analysis here this
is manageable, but a general-purpose accessibility ranking would need a
more careful POI selection methodology.

```{r create-location-column}
berlin_pois <- berlin_pois |> mutate(
  loc = st_coordinates(geometry),
  coord = map2(loc[,2], loc[,1], ~c(.x, .y))
) |> select(-loc) |> as.data.frame()

berlin_pois |> glimpse()
```

Here we set up the function to get the total number of user ratings for
a given POI using the Google Places API. The function first performs a
text search to find the place ID based on the name.

```{r poi-add-reviews}
dotenv::load_dot_env() 
api_key = Sys.getenv("GOOGLE_API_KEY")

library(googleway)

get_ratings_total <- function(place_name, coord, api_key) {
  res <- google_places(
    search_string = place_name,
    location = coord,
    key = api_key
  )

  res <- google_place_details(
    place_id = res$results$place_id[1],
    key = api_key
  )
  if( "user_ratings_total" %in% names(res$result) ) {
    res$result$user_ratings_total
  } else {
    0
  }
}
```

Let's test the function on a few known POIs to confirm it's working
correctly before applying it to the entire dataset. We can compare the
results to the Google Maps app for validation.

```{r poi-test-reviews}
berlin_pois |>
  filter(name %in%
    c(
      "Friedrichstadt-Palast",
      "Quatsch Comedy Club",
      "Nordamerikanischer Baumstachler"
    )) |>
  rowwise() |>
  mutate(
    user_ratings_total =
      tryCatch(
        get_ratings_total(name, coord, api_key),
        error = function(e) {
          NA
        }
      )
  ) |>
  ungroup()
```

The result matches the Google Maps app, confirming that the geocoding
and place.id resolution are working correctly.

::: callout-warning
Executing the next chunk of code may take a few minutes due to the
number of API calls and the 100ms sleep between them to respect rate
limits. It will fall under the free tier of 10,000 calls of Google Maps
API, but if you re-run it multiple times or scale up the number of POIs
or hotels, you may incur costs. Consider caching results or using a
smaller sample for testing. Google Maps pricing details can be found
[here](https://developers.google.com/maps/billing-and-pricing/pricing#places-pricing).
:::

The `tryCatch` ensures that if any individual POI lookup fails (due to
rate limits, geocoding issues, or network errors), the loop continues
and the failed POI gets an `NA` for its `user_ratings_total`, which can
be filtered out later.

The resulting `berlin_pois` dataframe will have a new column
`user_ratings_total` with the total number of user ratings for each POI,
which can be used for filtering and weighting in the subsequent
analysis.

```{r poi-filter-reviews}
#| eval: false
berlin_pois <- berlin_pois |>
  rowwise() |>
  mutate(
    user_ratings_total =
      tryCatch(
        get_ratings_total(name, coord, api_key),
        error = function(e) { NA }
      )
  ) |> 
  ungroup()
```

Once we have the ratings data, we save the enriched POI dataset to disk.
This allows us to avoid re-running the API calls in future sessions,
which is useful given the time and cost associated with those calls.

```{r save-pois-with-ratings}
#| eval: false
berlin_pois |> saveRDS("data/berlin_pois_with_ratings.rds")
```

The log scale on the histogram below is diagnostic; it shows the shape
of the distribution without compressing the long tail into invisibility.
The filtering itself still operates on the raw count.

```{r fig-poi-ratings-histogram}
berlin_pois <- readRDS("data/berlin_pois_with_ratings.rds")

ggplot(berlin_pois, aes(x = log1p(user_ratings_total))) +
  geom_histogram(fill = "lightblue", color = "black") +
  labs(
    title = "Distribution of User Ratings Total for POIs",
    x = "User Ratings Total",
    y = "Count"
  ) +
  scale_x_continuous(
    breaks = scales::pretty_breaks(n = 10),
    labels = function(x) round(expm1(x))
  ) +
  theme_minimal()
```

## Select Popular POIs

```{r poi-filter-top}
berlin_pois <- berlin_pois |>
  sort_by(desc(berlin_pois$user_ratings_total))

berlin_pois |>
  head(10) |>
  select(name, fclass, user_ratings_total) |> kable() 
```

*What is "Tresorfabrik Arnheim" by the way?* A quick Google search
suggests the Tresorfabrik S. J. Arnheim (founded in 1833) was the first
and largest manufacturer of safes and vaults in Germany, located in the
Berlin district of Wedding. Either most of those 40,160 Google reviews
are from the 19th century, or the ratings belong to a different
attraction. Let's exclude it from the results. We probably should have
started the search with the Google Maps API to avoid this kind of error.

Let's see how the top 50 POIs are distributed across classes.

```{r}
# Exclude Tresorfabrik Arnheim
berlin_pois <- berlin_pois |>
  filter(name != "Tresorfabrik Arnheim")
# Check class distribution of top 50 POIs
berlin_pois |>
  head(50) |>
  group_by(fclass) |> count()
```

```{r fig-popular-pois-map, message=FALSE}
library(ggplot2)
library(sf)
library(ggspatial)

berlin_pois$lng <- berlin_pois$coord |> map_dbl(2)
berlin_pois$lat <- berlin_pois$coord |> map_dbl(1)

berlin_pois_sf <- berlin_pois |> 
  head(50) |> 
  st_as_sf(coords = c("lng", "lat"), crs = 4326)

poi_shapes <- c(
  "attraction" = 17, 
  "museum"     = 15, 
  "theatre"    = 18, 
  "nightclub"  = 19, 
  "ruins"      = 8
)

poi_colors <- c(
  "attraction" = "#E41A1C",
  "museum"     = "#377EB8",
  "theatre"    = "#4DAF4A",
  "nightclub"  = "#984EA3",
  "ruins"      = "#FF7F00"
)

ggplot() +
  annotation_map_tile(type = "osm", zoom = 12) +
  geom_sf(data = st_transform(berlin_pois_sf, 3857), 
          aes(shape = fclass, color = fclass), 
          size = 3) +
  scale_shape_manual(values = poi_shapes, name = "Category") +
  scale_color_manual(values = poi_colors, name = "Category") +
  theme_minimal() +
  theme(
    axis.text = element_blank(),
    axis.title = element_blank(),
  ) +
  labs(
    title = "Berlin POIs by Category",
    subtitle = "Custom markers for cultural and nightlife locations"
  ) +
  theme(legend.position = "right")
```

The spatial distribution is well-suited for the analysis — POIs are not
concentrated purely at the centre, so the transit time signal will vary
meaningfully across the hotel map.

```{r filter-top-50-pois}
berlin_pois <- berlin_pois |> head(50)
```

## Collect Hotel Locations

```{r create-hotels}
hotels <- berlin_features |>
  filter(fclass == "hotel") |>
  select(osm_id, name, geometry)

hotels |> glimpse()
```

What about the spatial distribution of hotels? Are they clustered in a
few neighbourhoods, or spread evenly across the city? This will inform
the clustering strategy for routing.

```{r fig-hotels-map, message=FALSE}
library(ggplot2)
library(sf)
library(ggspatial)

hotels <- hotels |> mutate(
  loc = st_coordinates(geometry),
  coord = map2(loc[,2], loc[,1], ~c(.x, .y))
) |> select(-loc) |> as.data.frame()

hotels$lng <- hotels$coord |> map_dbl(2)
hotels$lat <- hotels$coord |> map_dbl(1)

hotels_sf <- hotels |> 
  st_as_sf(coords = c("lng", "lat"), crs = 4326)

ggplot() +
  annotation_map_tile(type = "osm", zoom = 12) +
  geom_sf(
    data = st_transform(hotels_sf, 3857), color = "red", size = 2
  ) +
  theme_minimal() +
  theme(
    axis.text = element_blank(),
    axis.title = element_blank(),
  ) +
  labs(
    title = "Berlin Hotels"
  )
```

455 hotels × 50 POIs = 22,750 routing requests. At Google Maps Distance
Matrix API pricing that's non-trivial, and the time cost of 100ms sleep
between calls makes it impractical to run in full. Clustering reduces
both cost and runtime. Failure handling is a separate concern — that's
what the `tryCatch` in the loop is for.

### Cluster Hotels with H3

We cluster hotels using H3 — Uber's hexagonal hierarchical geospatial
indexing system. Each hotel is assigned to its hexagonal cell at a given
resolution; hotels sharing a cell are represented by the cell centroid
for routing. The choice of hexagons over a regular square grid is
geometric: in a square cell, the distance from the centre to a corner is
$\sqrt{2}$ times the distance from the centre to an edge midpoint. In a
regular hexagon, the centre-to-edge distance is constant on all six
sides. For a routing approximation where the centroid stands in for
every point in the cell, that uniformity matters — the maximum
positional error is bounded and consistent, regardless of where within
the cell a hotel actually sits.

```{r h3-clustering, message=FALSE}
library(h3jsr)

hotels <- hotels |>
  mutate(
    h3_index = point_to_cell(geometry, res = 8)
  )

hotels_h3 <- hotels |>
  select(h3_index, geometry) |> 
  group_by(h3_index) |>
  summarise(
    hotel_count = n(),
    geometry = cell_to_polygon(h3_index)
  ) |> distinct()

hotels_h3 |> glimpse()
```

At resolution 8, cells have a mean area of approximately 0.74 km²,
reducing 455 hotels to 189 unique cells and cutting API calls by more
than half.

```{r fig-hotels-cluster-map, message=FALSE}
library(ggplot2)
library(sf)
library(ggspatial)

hotels_h3_sf <- st_as_sf(hotels_h3) |>
  st_transform(3857)

ggplot() +
  annotation_map_tile(type = "osm", zoom = 12) +
  geom_sf(
    data = hotels_h3_sf,
    aes(fill = hotel_count),
    color = "#333333",
    linewidth = 0.1,
    alpha = 0.8
  ) +
  scale_fill_viridis_c(
    option = "magma",
    name = "Hotel Count", direction = -1
  ) +
  theme_minimal() +
  theme(
    axis.title = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank()
  ) +
  labs(
    title = "Berlin Hotel Density",
    subtitle = "H3 Hexagons shaded by hotel count"
  )
```

189 origins × 50 destinations = 9,450 routing requests. Manageable.

```{r poi-group-centers}
hotels_h3 <- hotels_h3 |>
  mutate(
    center = h3jsr::cell_to_point(h3_index)
  )

hotels_h3 <- hotels_h3 |>
  rowwise() |>
  mutate(
    center = st_coordinates(center),
    lat = center[,2],
    lng = center[,1]
  ) |>
  select(-center) |>
  distinct()

hotels_h3 |> glimpse()
```

## Calculate Travel Times

```{r prepare-travel-time-calculation}
hotels_h3 <- hotels_h3 |>
  mutate(
    origin = paste0(lat, ", ", lng)
  )

origins <- hotels_h3 |>
  pull(origin) |>
  unique()

berlin_pois <- berlin_pois |>
  mutate(
    dest = paste0(
      st_coordinates(geometry)[, 2],
      ", ",
      st_coordinates(geometry)[, 1]
    )
  )

destinations <- berlin_pois |>
  head(50) |>
  pull(dest) |>
  unique()

paste("Number of origins:", length(origins))
paste("Number of destinations:", length(destinations))
```

The loop wraps each request in `tryCatch` as the Google Maps API
occasionally returns errors on rate limit spikes, ambiguous geocoding,
or network timeouts, and a bare `for` loop dies on the first one. The
results are collected into a pre-allocated list rather than grown with
`rbind`: `rbind` inside a loop copies the entire growing dataframe on
every iteration, an $O(n^2)$ memory pattern that is tolerable at 9,450
pairs but becomes the bottleneck if you scale to a larger city or expand
the POI set. `idx` is incremented outside the `tryCatch` so failed pairs
leave a `NULL` in the list; `bind_rows` drops NULLs silently.

::: callout-warning
The next code chunk makes 9,450 API calls to Google Maps. This may take
several minutes to run and could incur costs if you exceed the free tier
limits (see pricing info
[here](https://developers.google.com/maps/billing-and-pricing/pricing#routes-pricing)).
Consider caching results or using a smaller sample for testing.
:::

```{r compute-travel-times}
#| eval: false
library(ggmap)
library(gmapsdistance)

set.api.key(api_key)

# Pre-allocate a list to avoid rbind() inside the loop.
# rbind() copies the growing dataframe on every iteration — O(n²) memory
# behaviour that becomes noticeable above ~5k rows.
n_pairs <- length(origins) * length(destinations)
results_list <- vector("list", n_pairs)
idx <- 1

for (origin in origins) {
  for (destination in destinations) {
    tryCatch(
      {
        Sys.sleep(0.1)
        result <- gmapsdistance(
          origin      = origin,
          destination = destination,
          mode        = "transit"
        )
        results_list[[idx]] <- data.frame(
          origin      = origin,
          destination = destination,
          distance    = result$Distance,
          time        = result$Time
        )
      },
      error = function(e) {
        message(paste(
          "Error calculating travel time from", origin,
          "to", destination, ":", e$message
        ))
      }
    )
    idx <- idx + 1  # outside tryCatch: failed pairs leave NULL, 
                    # bind_rows drops them
  }
}

# Single bind at the end — fast regardless of row count
travel_times <- dplyr::bind_rows(results_list)

travel_times |> write.csv(
  "data/berlin_hotel_poi_travel_times.csv",
  row.names = FALSE
)
```

```{r fig-travel-times-histogram}
#| fig-cap: "Distribution of Travel Times from Hotels to POIs"
travel_times <- read.csv("data/berlin_hotel_poi_travel_times.csv")

ggplot(travel_times, aes(x = time / 60)) +
  geom_histogram(fill = "lightblue", color = "black", binwidth = 5) +
  labs(
    title = "Distribution of Travel Times from Hotels to POIs",
    x = "Travel Time (minutes)",
    y = "Count"
  ) +
  theme_minimal()
```

## Merge Travel Times with Hotel Clusters

```{r join-travel-times-with-hotels}
travel_times <- read.csv("data/berlin_hotel_poi_travel_times.csv")

hotels_h3_joined <- hotels_h3 |>
  left_join(
    travel_times,
    by = "origin"
  )

hotels_h3_joined |> saveRDS("data/berlin_hotels_with_travel_times.rds")
```

## Scenarios

The three scenarios model different traveller preference profiles: a
weighted random sample across all POI classes, museums and theatres
only, and attractions and nightclubs only. Hotel rankings change
substantially between scenarios. A hotel in Prenzlauer Berg performs
well for nightlife (transit spread across the city) but mediocrely for
museums (which cluster in Mitte, favouring more central locations).
Charlottenburg shows the opposite pattern.

```{r scenario-1}
hotels_h3_joined <- readRDS("data/berlin_hotels_with_travel_times.rds")

# Scenario 1: Weighted random sample of 7 POIs across all classes
set.seed(123)
preferred_pois <- berlin_pois |>
  sample_n(7, weight = user_ratings_total) |> select(dest) 
```

The visualisation uses inverse distance weighting (IDW) interpolation
over hex centroids to produce a continuous travel time surface, then
draws isolines at 5-minute intervals. IDW assumes that accessibility at
any unsampled point is a weighted average of the sampled hex centroids
around it, with weight falling off as a power of distance. It doesn't
extrapolate well beyond the data boundary, which is why the interpolated
surface is clipped to the convex hull of the hotel cluster before
plotting — no phantom isolines in unpopulated areas.

A note on the grid resolution: `dx = 0.005` (approximately 500m at
Berlin's latitude) is a deliberate tradeoff — finer grids improve
isoline detail but increase IDW computation time roughly quadratically.
At `dx = 0.001` the contours are noticeably smoother; at `dx = 0.005`
they run in acceptable time for an exploratory analysis.

The IDW grid is created in EPSG:4326 (degrees), then transformed to
EPSG:3857 (metres) to align with the OSM base tile from
`annotation_map_tile`.

The red circles at 1 km, 3 km, and 5 km from the geographic centre make
the core argument visual: some of the best-performing locations in
Scenario 3 sit well outside the 5 km ring.

```{r fig-top-hotels-map, message=FALSE}
library(ggplot2)
library(sf)
library(ggspatial)
library(gstat)
library(stars)
library(metR)
library(dplyr)

plot_sample_map <- function(sampled_pois) {
  # 1. Data Preparation
  hotels_h3_sample <- hotels_h3_joined |>
    inner_join(sampled_pois, by = c("destination" = "dest")) |>
    group_by(h3_index, hotel_count, geometry) |>
    summarise(avg_travel_time = mean(time), .groups = "drop") |>
    filter(avg_travel_time < quantile(avg_travel_time, 0.9)) |>
    st_as_sf()

  poi_data <- berlin_pois |>
    inner_join(sampled_pois, by = "dest") |>
    st_as_sf()

  # 2. Interpolation for Isolines
  hotel_hull <- st_convex_hull(st_union(hotels_h3_sample))
  grid <- hotels_h3_sample |>
    st_bbox() |>
    st_as_stars(dx = 0.005, dy = 0.005) |>
    st_set_crs(4326)

  idw_df <- idw((avg_travel_time / 60) ~ 1, hotels_h3_sample, grid) |>
    st_as_sf(as_points = TRUE) |>
    st_filter(hotel_hull) |>
    st_transform(3857)

  idw_df$x <- st_coordinates(idw_df)[, 1]
  idw_df$y <- st_coordinates(idw_df)[, 2]
  idw_df <- as.data.frame(idw_df)

  # 3. Reference Geometry (1/3/5 km circles from geographic centre)
  berlin_center <- st_sfc(st_point(c(13.4050, 52.5200)), crs = 4326)
  circle_1km <- berlin_center |>
    st_transform(3035) |>
    st_buffer(1000) |>
    st_transform(3857)
  circle_3km <- berlin_center |>
    st_transform(3035) |>
    st_buffer(3000) |>
    st_transform(3857)
  circle_5km <- berlin_center |>
    st_transform(3035) |>
    st_buffer(5000) |>
    st_transform(3857)

  # 4. Plotting
  ggplot() +
    annotation_map_tile(type = "osm", zoom = 12, alpha = 0.7) +
    geom_sf(
      data = st_transform(hotels_h3_sample, 3857),
      aes(fill = avg_travel_time / 60),
      color = "#333333", linewidth = 0.1, alpha = 0.5
    ) +
    geom_contour(
      data = idw_df, aes(x = x, y = y, z = var1.pred),
      breaks = seq(0, 200, by = 5), color = "black", linewidth = 0.4
    ) +
    geom_text_contour(
      data = idw_df, aes(x = x, y = y, z = var1.pred),
      breaks = seq(0, 200, by = 5), stroke = 0.15, size = 2.5
    ) +
    geom_sf(
      data = st_transform(poi_data, 3857),
      aes(shape = fclass, color = fclass), size = 3
    ) +
    geom_sf(data = circle_1km, fill = NA, color = "red", linewidth = 0.5) +
    geom_sf(data = circle_3km, fill = NA, color = "red", linewidth = 0.5) +
    geom_sf(data = circle_5km, fill = NA, color = "red", linewidth = 0.5) +
    scale_fill_gradientn(
      colors = c("green", "yellow", "red"),
      name = "Time (min)"
    ) +
    scale_shape_manual(values = c(
      "attraction" = 17, "museum" = 15, "theatre" = 18,
      "nightclub" = 19, "ruins" = 8
    ), name = "POI Type") +
    scale_color_manual(values = c(
      "attraction" = "#E41A1C", "museum" = "#377EB8",
      "theatre" = "#4DAF4A", "nightclub" = "#984EA3", "ruins" = "#FF7F00"
    ), name = "POI Type") +
    guides(
      fill = guide_colorbar(order = 1),
      shape = guide_legend(order = 2),
      color = guide_legend(order = 2)
    ) +
    coord_sf(crs = 3857) +
    theme_minimal() +
    theme(
      axis.text = element_blank(),
      axis.title = element_blank()
    ) +
    labs(
      title = "Berlin Accessibility & Cultural POIs",
      subtitle = "Shaded hexagons with 5-minute travel time contours"
    )
}

plot_sample_map(preferred_pois)
```

```{r scenario-2}
# Scenario 2: Focus on museums and theaters
preferred_pois <- berlin_pois |>
  filter(fclass %in% c("museum", "theatre")) |>
  sample_n(7, weight = user_ratings_total) |> select(dest)

plot_sample_map(preferred_pois)
```

```{r scenario-3}
# Scenario 3: Focus on nightclubs and attractions
preferred_pois <- berlin_pois |>
  filter(fclass %in% c("attraction", "nightclub")) |>
  sample_n(7, weight = user_ratings_total) |> select(dest)

plot_sample_map(preferred_pois)
```

## What's Next

The proof of concept demonstrates the method. A production version would
need:

-   **Dynamic POI selection** — let the user specify their points of
    interest rather than selecting from a fixed list of 50.

-   **Departure time sensitivity** — transit times in Berlin vary
    significantly between peak and off-peak hours; a morning departure
    to a museum is a different routing problem than a late-night return
    from a club.

-   **Kano validation** — conduct a research to validate the perceived
    value, see [Kano Method for Prioritization of
    Features](https://frequentist.org/posts/20240805-kano-model/).

The full code is in the [GitHub
repository](https://github.com/AxesAccess/H3-Routing-Isoline-R).
The processed travel time data is included for reproducibility without
re-running the API calls.

------------------------------------------------------------------------

**Author**

I'm Aleksei Prishchepo, an independent BI and analytics consultant based
in Berlin. I work on the analytical solutions: pipelines, models, and
dashboards with a focus on revenue analytics, geospatial analysis, and
marketing mix modelling. I write about practical data engineering and
analytics at [frequentist.org](https://frequentist.org). Contact me at: [linkedin.com/in/aleksei-pr](https://www.linkedin.com/in/aleksei-pr/).
